---
title: "R 101"
author: "Jinliang Yang"
date: "05-17-2019"
output:
  word_document: default
  pdf_document: default
  html_notebook: default
---

My first R code

```{r setup}
# , include=TRUE, warning=FALSE, echo=TRUE, error=FALSE
knitr::opts_knit$set(root.dir=normalizePath('../../'))
knitr::opts_chunk$set(warning=FALSE, message=FALSE, error=FALSE, echo=TRUE)
```


My first R code

```{r}
# Creating a Graph
attach(mtcars)
plot(wt, mpg) 
abline(lm(mpg~wt))
title("Regression of MPG on Weight")

# first row contains variable names, comma is separator 
# assign the variable id to row names
# note the / instead of \ on mswindows systems 

mydata <- read.table("C:/Users/Eric/Desktop/labworkshop/data/yu_etal_2015_SRR.csv", header=TRUE, sep=",")
head(mydata)
tail(mydata)
dim(mydata)
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


How to import data

```{r}
#read csv
Sra<- read.table("C:/Users/Eric/Desktop/labworkshop/data/SraRunTable_walley_etal_2016.txt", header=TRUE, 
   sep="\t")

head(Sra)#take a look at the first and last 6 rows
tail(Sra)
dim(Sra)




```
