---
title: "Tutorial about raw phenotypic data processing"
author: "Jinliang Yang"
date: "06-21-2019"
output:
  html_document:
    df_print: paged
---


```{r setup, include=TRUE}
#setwd("~/Documents/Github/pvpDiallel/")
knitr::opts_knit$set(root.dir=normalizePath('../../'))
```

Resave the data obtained from Semra `organized phenotyping data.xlsx` to `organized_phenotyping_data_Havelock2018.csv`  
- remove the special characters  


```{r, eval=FALSE}
# use this one to find out the special character:
# raw <- read.csv("data/organized_phenotyping_data_Havelock2018.csv", header=TRUE, stringsAsFactors=FALSE, fileEncoding="latin1")

raw <- read.csv("data/organized_phenotyping_data_Havelock2018.csv")
head(raw)
tail(raw)
dim(raw)
names(raw)
```

# Data sanity check (quality check)

## get cob width data only

```{r}
cw <- raw[,c("Barcode", "Number.of.ears", "Cob.width.1", "Cob.width.2", "Cob.width.3", "Cob.width.4",
                     "Cob.width.5", "Cob.width.6", "Cob.Width.7", "Cob.width.8")]
head(cw)
```

### Is barcode unique?

```{r}
dim(cw)
# count the length of non-unique barcode
length(unique(cw$Barcode))

# find out which one is not unique
idx <- which(duplicated(cw$Barcode))
# show the duplicated rows
cw[cw$Barcode %in% cw$Barcode[idx], ]

# for now, let's remove both rows
cw1 <- cw[!(cw$Barcode %in% cw$Barcode[idx]), ]
```



### data transformation

```{r}
# reshape the data from wide format to long format for cob idth and cob length trait
library("tidyr")
class(cw1[,3])

## important trick!
cw1[,3] <- as.numeric(as.character(cw1[,3]))

cw2 <- gather(cw1, key="Trait", value="value", c("Cob.width.1", "Cob.width.2", "Cob.width.3", "Cob.width.4",
                                                    "Cob.width.5", "Cob.width.6", "Cob.Width.7", "Cob.width.8"))

```

### summarize from the long format

```{r}
# remove the NA rows
cw2 <- subset(cw2, !is.na(value))
# order the data.frame by "Barcode" column
cw2 <- cw2[order(cw2$Barcode),]

cw2$Barcode <- as.character(cw2$Barcode)
# calculate for each bag "barcode", how many non-NA records
tb <- as.data.frame(table(cw2$Barcode)) #1210
# in the raw data 

tb <- merge(tb, cw1[, 1:2], by.x="Var1", by.y="Barcode", all=TRUE) #1210
#subset(tb, is.na(Number.of.ears))
subset(tb, is.na(Freq))
subset(cw1, Barcode %in% "2018-2477OP_DE811OP")

c1 <- subset(tb, Freq != Number.of.ears)
# 60 of the records are inconsistent with the obs.
subset(cw1, Barcode %in% "2018-1039OP_Tzi18OP")

```

### outlier by histogram

```{r}

median(cw2$value) #22.2
mean(cw2$value) #20.69717
range(cw2$value) #1.3 522.0

hist(cw2$value)
hist(cw2$value, breaks=100)
# https://www.color-hex.com/color-names.html
hist(cw2$value, breaks=100, col="#eec591", xlab="Cob Width", ylab="freq", main="histogram")


```

```{r}
cw2 <- cw2[order(cw2$value, decreasing=TRUE),]
cw2 <- cw2[-1:-2,]

hist(cw2$value, breaks=100, col="#eec591", xlab="Cob Width", ylab="freq", main="histogram")

```

### remove these extreme small values

```{r}
cw3 <- subset(cw2, value > 5)
hist(cw3$value, breaks=50, col="#eec591", xlab="Cob Width", ylab="freq", main="histogram")

```

### Is this a normal distribution?

```{r}
shapiro.test(cw3$value)

### what a normal distribution looks like
n <- rnorm(1000, mean = 5, sd = 3)
hist(n, breaks=50, col="#eec591", xlab="Cob Width", ylab="freq", main="histogram")

shapiro.test(n)
```



---------------------------

# More close to biology

## Seperate OP and @

```{r}
cw3$type <- "self"
idx <- grep("OP", cw3$Barcode)
cw3$type <- as.character(cw3$type)
cw3$type[idx] <- "OP"

table(cw3$type)
```

## Focus on OP lines

[Regular Expression Syntax:](http://www.endmemo.com/program/R/gsub.php)

```{r}
op <- subset(cw3, type == "OP")

  
## pattern match
op$row <- as.character(op$Barcode)
op$row <- gsub("_.*", "", op$row)
op$row <- gsub(".*-|OP", "", op$row)

op$row <- as.numeric(op$row)

op$row <- gsub(".*-|OP|op", "", op$row)
op$row <- as.numeric(op$row)
```


## assign plot id

```{r}
op$plot <- 0
op[op$row < 2000, ]$plot <- "plot1"
op[op$row > 2000 & op$row < 3000, ]$plot <- "plot2"
op[op$row > 3000 & op$row < 4000, ]$plot <- "plot3"
op[op$row > 4000, ]$plot <- "plot4"

table(op$plot)
```

# histogram using ggplot2 

[quickR](http://www.sthda.com/english/wiki/ggplot2-histogram-plot-quick-start-guide-r-software-and-data-visualization)

```{r}
library("ggplot2")
p<-ggplot(op, aes(x=value, fill=plot)) + 
  geom_histogram()
p

```



## Use facets

```{r}
p<-ggplot(op, aes(x=value))+
  geom_histogram(color="black", fill="white")+
  facet_grid(plot ~ .)
p
```




















